---
title: "San Fransciso Trees"
author: 'null'
date: '2020-12-08'
output: pdf_document
categories: []
tags: []
slug: san-fransciso-trees
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyverse)
library(here)
library(patchwork)
```

Using the sf_trees.csv dataset from Tidy Tuesday, I investigate what the trees
are like in San Francisco. The data has 192,987 observations and 12 variables. 
The variables are the ID given to the tree, the legal status of the tree, the 
species of the tree, the street address, the order of the trees in the case that 
there are multiple trees at one location, where the tree is on the lot, the 
primary caretaker of the tree, the date it was planted, the diameter of the tree, 
the dimension of the plot, the latitude the plant is at, and the longitude the 
tree is planted.

```{r}
sf <- read.csv(here("tidytuesday-master", "data", "2020", "2020-01-28", "sf_trees.csv"))
```

# Question 1 : Where are the trees in San Francisco and who owns them?

```{r}
ggplot(sf, aes(x = (longitude*-1), y = latitude, color = caretaker)) +
  geom_point(size = 0.001, alpha = 0.2) +
  ylim(37.7,37.825) +
  theme(legend.position = "none") +
  scale_x_continuous(trans = "reverse") +
  xlim(122.52, 122.36) +   
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Map of the Trees in San Francisco")
```

This reveals that there is a lot about where the trees are in San Francisco. A
person can easily see the street layout of San Francisco. Also, a person can see
how well the streets are lined with trees. It also shows where the data set 
didn't collect data, which is parks. The long strip in the left side of the map
is actually Golden Gate Park which has several trees, but these trees were not 
included in the scope of the research. The central blank area as well as the 
North West and South West are areas reserved for more nature, usually golf 
courses. the only area with no trees that actually has no trees is the area in 
the South East. This is a primarily commercial area with few street decorations 
or Trees. The blue indicates private trees, which suggests that while the trees 
may be mandatory to be on a street, it is still the land owner's responsibility 
to take care of the tree. 

# Question 2: What is the most common type of tree?

```{r}

species_list <- sf %>%
  mutate(species = word(species, 1, sep=" ")) %>%
  separate_rows(species, sep = ' ') %>%
  group_by(species) %>%
  summarize(Count = n()) %>%
  filter(Count >= 5000)

x <- 0

for (i in 1:12) {
  x <- x + species_list[i,2]
}

z <- 19287 - x

species_list[nrow(species_list) + 1,] = c("Other", z)

ggplot(species_list, aes(x= "", y=Count, fill = species)) +
  geom_bar(width = 10, stat = "identity") +
  coord_polar("y", start = 0) +
  xlab(" ") +
  ylab(" ") +
  ggtitle("Number of Trees of a Certain Species")

```

Surprisingly, there is a lot of variety in tree. About 50% of trees have less 
than 5000 of the same type of tree. Considering there are around 200,000 trees,
it means that only 11 tree types are over 2.5% of the trees in the San Francisco
area. This indicates that the people care about their yards and their plants and 
want their yards to be more unique and have personality, so they choose a more 
unique tree. The area also must be habitable for any different tree types because
there is so much of a variety. If there was a large potion of any one type of 
tree, it would be assumed that there is a smaller tree pool to choose from for 
consumers. However, it looks like people have found that many trees can grow 
well in the San Francisco area.

# Question 3: Does site of the tree affect the diameter of the tree?

```{r}
g1 <- ggplot(sf, aes(x= site_info)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90)) +
  xlab("Tree Location") +
  ylab("Number of Trees") +
  ggtitle("Popular Locations of Trees in SF")

g2 <- ggplot(sf, aes(x= site_info)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90)) + 
  ylim(0,12500) +
  xlab("Tree Location") +
  ylab("Number of Trees") +
  ggtitle("Locations Excluding Curbside Cutouts")

g1 + g2

```

As seen in the graph, a majority of the trees are on sidewalk curbs. This is 
most likely why we see such a distinct street layout in the spatial map. These 
trees most likely consists of the more common tree types because they are on 
streets. Since they are on streets, it suggests that there is some degree of 
uniformity set in place by some form of city official or home owners association.
Of the relatively smaller tree locations, they are mostly in yards. This I believe
is where the variation in tree types comes from. In yards, it is more likely that
people can take more creative freedom on what kind of tree they want. While some
trees look nice on a curb, other more exciting and unique trees are in private 
yards to add more character to one's home.

```{r}
today <- 2020

tree_age <- sf %>%
  filter(!is.na(dbh)) %>%
  mutate(age = substr(date, 1,4)) %>%
  mutate(age = as.numeric(age)) %>%
  mutate(age = today - age)
  
tree_age

b1 <- ggplot(tree_age, aes(x= site_info, y = dbh)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90)) +
  ylim(0,125) +
  xlab("Tree Location") +
  ylab("Diameter") +
  ggtitle("Diameter of Tree per Location")

b2 <- ggplot(tree_age, aes(x= site_info, y = age)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90)) +
  ylim(0,70) +
  xlab("Tree Location") +
  ylab("Age") +
  ggtitle("Age of Tree per Location")

b1 + b2

```

These graphs shows that trees that grow in yards grow bigger for their age then 
other trees. While other spots may use trees as decoration, trees should be in 
the place where they grow the best. 

# Conclusion

San Francisco has many beautiful trees in its cities that like to line the roads 
and street layout. It also has many trees that are in large parks as well as in 
people's personal yards. There is a lot of variation in the tree types, and no 
one tree has near a large portion at all. This adds to the diversity of the city
However, if trees grew in yards, the trees would grow bigger and better.
